---
title: "Usage of mscmposer on the CODE-DE platform"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usage of mscmposer on the CODE-DE platform}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r setup}
library(sf)
library(gdalcubes)
library(mscomposer)
```


This vignette is meant to show the how the package can be used on the [CODE-DE](https://code-de.org/)
platform. This platform requires a subscription plan in order to let you create 
virtual machines (VMs) on which you can use the `{mscomposer}` package.

Here, we assume that you signed-up for a subscription and have a Ubuntu based VM 
running with the CODE-DE data platform mounted. One easy way to achieve this during
the setup of your VM, is documented in the `{mscomposer}` repository. We explain 
there how to set up a docker container on your VM with `{mscomposer}` installed and
which is reachable via an R Studio interface through your web browser.

With this setup, the CODE-DE data platform will be mounted, both on the host VM 
and the docker container under the following path `/codede/`.
We can thus work with the Sentinel-2 data without the need to download to our local
machine. The first step consists of querying the EOFinder API, which Sentinel-2
scenes intersect with our area of interest (AOI). Remember, that `query_s2_codede()`
expects the aoi object to be in uprojected coordinates.

```{r eofinder}
aoi <- st_read(system.file("aoi_S2.gpkg", package = "mscomposer"))
(s2_matches <- query_s2_codede(
  aoi = aoi,
  level = "LEVEL2A",
  product = "L2A-MAJA",
  cloudcover = 10,
  start = "2022-01-01",
  end = "2022-12-31"
))

```

The `s2_matches` is an sf object with the spatial footprint of matching acquisitions
as the geometry and some metadata information as attributes. Assuming that
you are working on the CODE-DE platform, the column `productIdentifier` is of 
special interest, because this variable indicates the paths to the respective
acquisitions under `/codede/`. We can thus list all matching `*.tif` files for 
your composite processing.

```{r list-s2, eval = FALSE}
root_paths <- s2_matches$productIdentifier
s2_files <- list.files(root_paths, pattern = "*.tif", full.names = TRUE, recursive = TRUE)
```

Note, that we cannot actully run the above code when building this vignette, because
it will only find any files if run on a machine on CODE-DE. Given that assumption,
we than can proceed using `{ms_composer}` as laid out in the README, by firts 
constructing a view and then a composite based on the matched files and a specific
collection format:

```{r composer, eval = FALSE}

harz_view <- ms_view(
  aoi, 
  srs = st_crs(aoi_utm), 
  dt = "P30D", 
  dx = 20, 
  dy = 20, 
  start = "2022-01-01", 
  end = "2022-12-31", 
  agg = "mean", 
  rsmp = "bilinear"
)

ms_compose(
  files = s2_files,
  format= system.file("S2-L2A-MAJA-CODEDE.json", package = "mscomposer"),
  view = harz_view, 
  col = "harz_2022_maja.sqlite", 
  mask = image_mask("CLOUDS", min = 1, max = 255), 
  bands = c("B2", "B3", "B4", "B5", "B6", "B7", "B8", "B8A", "B11", "B12"),
  reducers = "median",
  indices = list(NDVI = "(B8-B4)/(B8+B4)"),
  filters = list(NDVI = "NDVI < 0.2"),
  outdir = ".",
  prefix = "harzvorland-composite-maja-"
)

```

In case that you do not have a subscription plan that allows you to create
a VM on CODE-DE, you may still create an account that will enable you to download
data to your local machine. `{mscomposer}` allows you to download matching 
Sentinel-2 acquisitions in case you provide it with an API key used to identify you 
as a subscribed user to CODE-DE. The below function call will download
the matching scenes to the current directory as zip files. You may wish to extract
the zip files after the download, or you may relay on GDAL's ability to read
files directly from zips using the `/vsizip/` driver. 

```{r download, eval = FALSE}
download_s2_codede(
  s2_matches,
  key = "<your-api-key>",
  outdir = "."
)

```



